Open-source tools in R for landscape ecology

### 2021-02-08 paper/paper.tex
add in LJG funding

The vector data model consists of two main elements i) geometries (such as points, lines, polygons) and ii) attributes, where each geometry is connected to a row in the attribute table.
In many cases, this data model allows a more realistic representation of landscape features, however, with the cost of higher computational demands \cite{Lovelace2019}.
The \emph{sp} package was the standard for vector data representation for more than ten years \cite{Pebesma2005,Bivand2013}.
As of 2020, more than 500 \emph{R} packages directly depend or imports \emph{sp}.
However, \emph{sp} is not actively developed anymore, and its recommended successor is the \emph{sf} package \cite{Pebesma2018}.
Besides many advantages and strength of the \emph{sf} package in terms of spatial data handling, it also integrates into the widely used \emph{tidyverse} packages \cite{Wickham2019}.
\emph{sf} builds on the idea of ``simple features,'' a standard used to describe spatial geometries using points, lines (two connected points) and polygons (several connected points) and attributes connected to these geometries \cite{Pebesma2019a}.

@@

We asked the participants to select which terms describe their research topics the best, and options that were selected by more than 10\% of participants included ``biodiversity,'' followed by ``land use management,'' ``landscape connectivity,'' and ``nature conservation'' (Fig. \ref{fig:fig-survey}a)).

Next, we were interested in the most important tasks to the workflow of the participants.
Not surprisingly, ``(pre-)processing of data,'' ``spatial statistics,'' and ``creating maps'' were the most selected options (Fig. \ref{fig:fig-survey}b).
The available options seemed to describe the most important task to the workflow quite well since only very few participants selected the ``others'' option (all options with less than five total answers were classified as ``others'').

@@

Lastly, when asked how useful \emph{R} is currently for landscape ecology, the vast majority of participants answered with either ``very useful'' or ``useful'' (summarized 90.83\%) and only very few participants evaluated \emph{R} as ``intermediate,'' ``not useful'' or ``not useful at all'' (summarized 9.17\%; Fig. \ref{fig:fig-survey}d)).

@@

MHKH was supported by the German Research Association (DFG) Research Training Group 1644 `Scaling Problems in Statistics,' grant number 152112243.

>>>>

The vector data model consists of two main elements i) geometries (such as points, lines, polygons) and ii) attributes, where each geometry is connected to a row in the attribute table.
In many cases, this data model allows a more realistic representation of landscape features, however, with the cost of higher computational demands \cite{Lovelace2019}.
The \emph{sp} package was the standard for vector data representation for more than ten years \cite{Pebesma2005,Bivand2013}.
As of 2020, more than 500 \emph{R} packages directly depend or imports \emph{sp}.
However, \emph{sp} is not actively developed anymore, and its recommended successor is the \emph{sf} package \cite{Pebesma2018}.
Besides many advantages and strength of the \emph{sf} package in terms of spatial data handling, it also integrates into the widely used \emph{tidyverse} packages \cite{Wickham2019}.
\emph{sf} builds on the idea of ``simple features'', a standard used to describe spatial geometries using points, lines (two connected points) and polygons (several connected points) and attributes connected to these geometries \cite{Pebesma2019a}.

@@

We asked the participants to select which terms describe their research topics the best, and options that were selected by more than 10\% of participants included ``biodiversity'', followed by ``land use management'', ``landscape connectivity'', and ``nature conservation'' (Fig. \ref{fig:fig-survey}a)).

Next, we were interested in the most important tasks to the workflow of the participants.
Not surprisingly, ``(pre-)processing of data'', ``spatial statistics'', and ``creating maps'' were the most selected options (Fig. \ref{fig:fig-survey}b).
The available options seemed to describe the most important task to the workflow quite well since only very few participants selected the ``others'' option (all options with less than five total answers were classified as ``others'').

@@

Lastly, when asked how useful \emph{R} is currently for landscape ecology, the vast majority of participants answered with either ``very useful'' or ``useful'' (summarized 90.83\%) and only very few participants evaluated \emph{R} as ``intermediate'', ``not useful'' or ``not useful at all'' (summarized 9.17\%; Fig. \ref{fig:fig-survey}d)).

@@

MHKH was supported by the German Research Association (DFG) Research Training Group 1644 `Scaling Problems in Statistics', grant number 152112243.

LJG was supported by the Natural Environment Research Council (NERC) grant \emph{A complex-systems approach to improve understanding of the biodiversity-landscape structure relationship}, grant number NE/T009373/1.
